Differences in Homicide Case Information Indicates Why Justice is Not Served

### 2024-12-14 paper/paper.qmd
Proofread Introduction and updated subtitle

---
title: "Differences in Homicide Case Information Indicates Why Justice is Not Served"
subtitle: "An analysis of solved and unsolved homicides from 2010 to 2017 in one of the United States's 2 largest cities, Chicago and Los Angeles"
author: 
  - Emily Su
thanks: "Code and data are available at: [https://github.com/moonsdust/unsolved-murders](https://github.com/moonsdust/unsolved-murders)."
date: today
date-format: long
abstract: "Despite Chicago Mayor Brandon Johnson announcing in 2024 that the Chicago Police Department had the highest homicide clearance rate of 54% in years, homicide clearance rates across the United States (US) have been decreasing since 1980, with more homicides going unsolved. This paper looks at patterns of homicide case information with the case resolution from 2010 to 2017 in two of the largest cities in the United States, Chicago and Los Angeles. We found that unsolved homicides are more likely to occur in Chicago compared to Los Angeles, and the majority of unsolved homicide victims are Black or Hispanic and are male. These findings can inform the public and US police departments of populations more vulnerable to having unsolved cases however further investigation is needed on the investigators involved in each unsolved homicide case."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

On November 20, 2024, Chicago Mayor Brandon Johnson announced that the Chicago Police Department had the highest homicide clearance rate, which are cases that are solved, as of recent years at almost 54% [@citechicagonews]. Despite celebrating this accomplishment, @citechicagonews found that if only the cases with arrest was considered and not closed cases where the police believed they have identified the perpetrator without an arrest being made, their clearance rate was actually 23% this year. This concern about homicide clearance rate was expressed by @citeguardian where they mentioned that homicide clearance rates from 1980 to 2020 have decreased from 71% to 50% in the United States (US). As of 2023, there are currently no up-to-date database keeping track of homicide resolutions and current crime databases do not indicate the victim's race and what led to the homicide, making it difficult to figure out why homicides are going unsolved [@citeguardian]. This raises the following question: what are the differences in homicide case information like the year and city the homicide took place and the victims’ perceived characteristics (age, sex, and race) between solved and unsolved homicides in one of the 2 of the largest cities in the United States, Chicago and Los Angeles, from 2010 to 2017?

@@

In this paper, to see what factors could impact the likelihood of a homicide case going unsolved, we analyzed data provided by The Washington Post [@datahomicides] that were compiled by their reporters on homicide cases and their status from 2010 to 2017. Previously, The Washington Post has used the data they compiled to construct a map and pinpointing places unsolved homicides occur [@citewashingtonpostmap]. In another study, @citeIndiana looked at factors that contributed to the low homicide clearance rates in Indianapolis, Indiana across different neighbourhoods. They found that there was an association between neighbourhoods with higher arrest rates and them having a higher chance of homicide clearance [@citeIndiana]. However, specific analysis looking at the differences in case characteristics between specific two US cities like Chicago and Los Angeles has not been done previously. In our findings, it indicates that a majority of unsolved homicide victims from 2010 to 2017 are Black or Hispanic and they are likely to be male but their age does not impact if there case goes unsolved or not. We also found that unsolved homicides are more likely to occur in Chicago compared to Los Angeles from 2010 to 2017. From our analysis, most unsolved homicide cases occurred during the middle and later half of the year with there being a jump in unsolved cases from 2016 onwards. These findings could inform the public and police departments about populations more vulnerable to homicide cases going unsolved and cities to be more alert about such as Chicago with homicide clearance rates.

For the remainder of the paper, the data section (@sec-data) goes over the data we used and its features and limitations, how the dataset was obtained, our variables of interest for our model and preliminary analysis, and a bit of our data cleaning process. In the model section (@sec-model), it discusses the type of model we used, the relationships between variables in our model, and the justification for our model. The results section (@sec-results) contains tables and graphs created based on our data and the results from our model. In the discussion section (@sec-discussion), it goes over what we did in our results, explaining the meaning of our findings, the implications of our findings to the real world, and areas of improvement and suggestions for future works. In the appendix (@sec-appendix), it includes a link to the dashboard with interactive versions of the graphs in this paper, additional tables, notes on reproducing and code styling, model diagnostics and validation, our idealized survey and methodology to gather more data on investigators responsible for homicide cases, and evaluation of The Washington Post's methodology.

@@

Some limitations with using a survey is that even though we designed it so that answers are required for all demographic questions and job-related questions and therefore minimize non-response errors [@citetellingstories], @citeStantcheva mentions that survey attrition becomes a problem. Attrition occurs when respondents decide to not complete the survey over the course of completing the survey [@citeStantcheva]. @citeStantcheva also noted respondent could have biases like social desirability bias and response order bias when responding to a survey. Our survey design aimed to mitigate these biases but they could still occur with respondents's answers.

>>>>

---
title: "Differences in Homicide Case Information Indicates Why Justice is Not Served"
subtitle: "An analysis of solved and unsolved homicides from 2010 to 2017 in one of the United States' 2 largest cities, Chicago and Los Angeles"
author: 
  - Emily Su
thanks: "Code and data are available at: [https://github.com/moonsdust/unsolved-murders](https://github.com/moonsdust/unsolved-murders)."
date: today
date-format: long
abstract: "Despite Chicago Mayor Brandon Johnson announcing in 2024 that the Chicago Police Department had the highest homicide clearance rate of 54% in years, homicide clearance rates across the United States (US) have been decreasing since 1980, with more homicides going unsolved. This paper looks at patterns of homicide case information with the case resolution from 2010 to 2017 in two of the largest cities in the United States, Chicago and Los Angeles. We found that unsolved homicides are more likely to occur in Chicago compared to Los Angeles, and the majority of unsolved homicide victims are Black or Hispanic and are male. These findings can inform the public and US police departments of populations more vulnerable to having unsolved cases however further investigation is needed on the investigators involved in each unsolved homicide case."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

On November 20, 2024, Chicago Mayor Brandon Johnson announced that the Chicago Police Department had the highest homicide clearance rate of 54% in recent years [@citechicagonews]. Despite celebrating this accomplishment, @citechicagonews found that if only the cases with arrest were considered and not closed cases where the police believed they had identified the perpetrator without an arrest being made, their clearance rate was instead 23% this year. This concern about homicide clearance rate was expressed by @citeguardian, who mentioned that homicide clearance rates from 1980 to 2020 have decreased from 71% to 50% in the United States (US). As of 2023, there are currently no up-to-date databases keeping track of homicide resolutions, and current crime databases do not indicate a victim's race and what led to a homicide, making it difficult to figure out why homicides are going unsolved [@citeguardian]. The ongoing concern about homicide clearance raises the following question: what are the differences in homicide case information like the year and city where the homicide took place and the victim's perceived characteristics (age, sex, and race) between solved and unsolved homicides in the 2 of the largest cities in the United States, Chicago and Los Angeles, from 2010 to 2017?

@@

In this paper, to see what factors could impact the likelihood of a homicide case going unsolved, we analyzed data provided by The Washington Post [@datahomicides] that was compiled by their reporters on homicide cases and their status from 2010 to 2017. Previously, The Washington Post used the data they assembled to construct a map and pinpoint places where unsolved homicides occur [@citewashingtonpostmap]. In another study, @citeIndiana looked at factors that contributed to the low homicide clearance rates in Indianapolis, Indiana across different neighbourhoods. They found that there was an association between neighbourhoods with higher arrest rates and having a higher chance of homicide clearance [@citeIndiana]. However, specific analysis looking at the differences in case characteristics between two US cities, Chicago and Los Angeles, has not been done previously. Our findings indicate that a majority of unsolved homicide victims from 2010 to 2017 are Black or Hispanic and are likely to be male but their age does not impact if their case goes unsolved or not. We also found that unsolved homicides are more likely to occur in Chicago compared to Los Angeles from 2010 to 2017. From our analysis, most unsolved homicide cases occurred during the middle and later half of the year with there being a jump in unsolved cases from 2016 onwards. These findings could inform the public and police departments about populations more vulnerable to homicide cases going unsolved and cities to be more alert about such as Chicago with homicide clearance rates.

For the remainder of the paper, the data section (@sec-data) goes over the data we used and its features and limitations, how the data was obtained, our variables of interest for our model and preliminary analysis, and a bit of our data cleaning process. The model section (@sec-model) discusses the type of model we used, the relationships between variables in our model, and the justification for our model. The results section (@sec-results) contains tables and graphs created based on our data and the results from our model. The discussion section (@sec-discussion) goes over what we did in our results, explaining the meaning of our findings, the implications of our findings to the real world, and areas of improvement and suggestions for future works. The appendix (@sec-appendix) includes a link to a dashboard that was created with interactive versions of the graphs in this paper, additional tables, notes on reproducing and code styling, model diagnostics and validation, our idealized survey and methodology to gather data on investigators responsible for homicide cases, and our evaluation of The Washington Post's methodology.

@@

Some limitations with using a survey is that even though we designed it so that answers are required for all demographic questions and job-related questions and therefore minimize non-response errors [@citetellingstories], @citeStantcheva mentions that survey attrition becomes a problem. Attrition occurs when respondents decide to not complete the survey over the course of completing the survey [@citeStantcheva]. @citeStantcheva also noted respondent could have biases like social desirability bias and response order bias when responding to a survey. Our survey design aimed to mitigate these biases but they could still occur with respondents' answers.
